\propertitle

### 2019-04-23 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-04-23T10:42:47+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation constraints%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample constraints}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The equality above extends to other quantities beyond the means $S/N$ and
$s/n$: it holds for all normalized \emph{factorial} moments
\citep{potts1953,broca2005}:
\begin{equation}
  \label{eq:equal_factorialmoments_polls}
  \begin{split}
  \E\condt[\Big]{\tfrac{s\;(s-1)}{n\;(n-1)}}{\yH} &= 
  \E\condt[\Big]{\tfrac{S\;(S-1)}{N\;(N-1)}}{\yH},\\
  \E\condt[\Big]{\tfrac{s\;(s-1)\;(s-2)}{n\;(n-1)\;(n-2)}}{\yH} &= 
  \E\condt[\Big]{\tfrac{S\;(S-1)\;(S-2)}{N\;(N-1)\;(N-2)}}{\yH},\quad\dotsc
  \\
  \text{and in general}\quad
  \E\condt[\Big]{\tbinom{s}{r}\;\tbinom{n}{r}^{-1}}{\yH} &=
  \E\condt[\Big]{\tbinom{S}{r}\;\tbinom{N}{r}^{-1}}{\yH},
  \quad r<N.
\end{split}
\end{equation}
\end{subequations}
Note that from the first $r$ factorial moments we can calculate the first
$r$ power moments and vice versa; but the simple equalities above don't
hold for the power moments \citep[\sect~A]{portamanaetal2015}.

The relations above are powerful because they hold no matter what the
probability distribution $\pf(S\|\yH)$ be.

@@

Whenever we record the activity of $n$ neurons we're effectively polling
them on the activity of all $N$ neurons under the recording device -- those
that could equally likely have been sampled instead. Suppose their
activities are time-binned and binarized. At every time bin $t$, $s(t)$ out
of the $n$ neurons \enquote{vote} for activity, and $n-s(t)$ vote for
silence. The probability calculus tells us that the expectation
formulae~\eqref{eq:all_equal_expectations_poll} hold for the total activity
$S(t)$ of the full population of $N$ neurons that could have been recorded.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-04-23T11:28:39+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation constraints%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample constraints}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The equality above extends to other quantities beyond the means $S/N$ and
$s/n$: it holds for all normalized \emph{factorial} moments
\citep{potts1953,broca2005}:
\begin{equation}
  \label{eq:equal_factorialmoments_polls}
  \begin{split}
  \E\condt[\big]{\tfrac{s\;(s-1)}{n\;(n-1)}}{\yH} &= 
  \E\condt[\big]{\tfrac{S\;(S-1)}{N\;(N-1)}}{\yH},\\
  \E\condt[\big]{\tfrac{s\;(s-1)\;(s-2)}{n\;(n-1)\;(n-2)}}{\yH} &= 
  \E\condt[\big]{\tfrac{S\;(S-1)\;(S-2)}{N\;(N-1)\;(N-2)}}{\yH},
  \\
  \text{and in general}\quad
  \E\condt[\big]{\tbinom{s}{r}\;\tbinom{n}{r}^{-1}}{\yH} &=
  \E\condt[\big]{\tbinom{S}{r}\;\tbinom{N}{r}^{-1}}{\yH},
  \quad r<N.
\end{split}
\end{equation}
Note that from the first $r$ factorial moments we can calculate the first
$r$ power moments and vice versa; but the simple equalities above don't
hold for the power moments \citep[\sect~A]{portamanaetal2015}. The formulae
above follow from a precise relation between our \dobs\ about the polled
voters and about all voters:
\begin{equation}
  \label{eq:relation_ps_pS}
  \pf(s \| \yH) = \sum_{S} \binom{n}{s}\;\binom{N-n}{S-s}\;\binom{N}{S}^{-1}\;
  \pf(S \| \yH).
\end{equation}
\end{subequations}
The relations above are powerful because they hold no matter what the
probability distribution $\pf(S\|\yH)$ may be.

@@

This probabilistic fact is of outstanding importance in neuroscience, where
we try to understand the workings of brain areas that may include tens of
thousands of neurons, but we can record only few tens or hundreds of them.
Whenever we record the activity of $n$ neurons we're effectively polling
them on the activity of all $N$ neurons under the recording device -- those
that could equally likely have been sampled instead. Suppose the recorded
activities are time-binned and binarized. At every time bin $t$, $s(t)$ out
of the $n$ neurons \enquote{vote} for activity, and $n-s(t)$ vote for
silence. The probability calculus tells us that the expectation
formulae~\eqref{eq:all_equal_expectations_poll} hold for the total activity
$S(t)$ of the full population of $N$ neurons that could have been recorded.

The simple expectation equalities are

Expectations are often used with
